1.6: Optimize Workflow: Projects, Markdown, and Git

### 2019-09-12 static/slides/06-slides.Rmd
Fix 06 slides

### [dplyr::select(): select variables](#60)

### [dplyr::rename(): rename variables](#73)

### [dplyr::mutate(): create new variables](#77)

### [dplyr::summarize(): create statistics](#98)

### [tidyr: reshaping data](#133)

### [dplyr: combining datasets](#134)

@@

![From [R Studio's R Markdown Cheatsheet](https://www.rstudio.com/wp-content/uploads/2016/03/rmarkdown-cheatsheet-2.0.pdf)](images/rmd2.png)

@@

.pull-right[
.center[
![](https://www.dropbox.com/s/czklhgt5llrekjc/newrmarkdown.png?raw=1)
]
]

@@

# Structure of an `R Markdown` Document

@@

3. `R` chunks for data analysis, plots, figures, tables, statistics, as necessary

.footnote[<sup>.red[1]</sup> The one exception is for managing bibliographies, this requires an additional `.bib` file!]

@@

- You can add [additional options](this guide](https://yihui.name/knitr/options/)) inside the `{braces}` after `r`, some common options:

@@

- [Markdown](https://en.wikipedia.org/wiki/Markdown) is a lightweight markup language geared towards HTML (i.e. the internet)
    - [Markup languages](https://en.wikipedia.org/wiki/Markup_language) used to add commands about how to display plain text
- Very simple and intuitive
- Write normal text as usual in any word processor
- Change font styling with tags (asterisks): 
    - `*italics text*` creates *italics text*
    - `**bold text**` creates **bold text**

@@

- A great resource: [Wikibooks LaTeX Mathematics chapter](https://en.wikibooks.org/wiki/LaTeX/Mathematics)
- And for more about $\LaTeX{}$ in general: [Wikibooks LaTeX](https://en.wikibooks.org/wiki/LaTeX/)

@@

- First term after the `{` is your **citation key** (e.g. `safner2016`)
]
---

@@

- Whenever you want to cite a work in your text, call up the **citation key** with `@`, like so: `@safner2016[]`, which produces @safner2016[]
    - You can customize citations, e.g.:
        - **All in parentheses**: `[@safner2016]` produces [@safner2016]
        - **No parentheses**: `@safner2016` produces @safner2016
        - **Date only**: `[-@safner2016]` produces [-@safner2016]
        - **Adding pages or comments**: `@safner2016[p.743-744 and passim]` produces @safner2016[p.743-744 and passim]
- BibTeX will automatically collect all works cited at the end and produce a **bibliography** according to a style you can choose

@@

- Markdown files are [**plain text**](https://en.wikipedia.org/wiki/Plain_text) files and can be edited in *any* text editor
    - something as basic (and boring!) as "**Notepad**," for example
    - many good [text editors](https://en.wikipedia.org/wiki/Text_editor) out there

@@

.footnote[Big shout outs also to [Typora](https://www.typora.io/) and [Ulysses](https://ulysses.app/) (Mac only) that both write (and preview) Markdown in a simple interface, no distractions.]

@@

- Honestly, I write **everything** in R Studio's text editor
    - Syntax highlighting
    - Actually can *run* R code, autocomplete, etc<sup>.red[1]</sup>
    - Can render the markdown to an output format: html, pdf, etc.

.footnote[<sup>.red[1]</sup> You can *write* R code in other text editors, but you can't *execute* them outside of *R Studio* (or the command line, but that's too advanced.) Same with actually rendering your markdown to an output (pdf, html, etc)]

@@

- When you are ready, you "compile" your markdown and code into an output format using: 
    - [`knitr`](https://yihui.name/knitr/)<sup>.red[1]</sup>, an R package that "`knit`s" your R code and markdown `.Rmd` into a `.md` file for:  
  - [pandoc](http://pandoc.org) is a "swiss-army knife" utility that can convert between *dozens* of document types

- All you need to do is click the `Knit` button at the top of the text editor!

.footnote[<sup>.red[1]</sup> `knitr` also relies on the `rmarkdown` package, which will probably be installed when you first `knit.`]
---

# knitr: What's Happening On Your Computer

.center[
![](https://www.dropbox.com/s/kzvb8n8fpf6lstz/pandoc1.png?raw=1)
]

@@

.center[
![:scale 80%](https://www.dropbox.com/s/i1ijatt56ncxi0e/github%202.png?raw=1)

- This project is on [GitHub](http://github.com/ryansafner/workflow)
- Click the green button, download to your computer, open `.Rproj` file in R Studio
]

@@

- `GitHub` is a popular (not the only!) cloud destination for these repositories

]
]
.pull-right[
- Shows history (`versions`) of files with comments
    - Can `fork` or `branch` repository into multiple versions at once
    - Good for "testing" things out without destroying old versions!
    - `revert` back to original versions as needed

@@

.center[
![:scale 70%](https://www.dropbox.com/s/egfojonn2s54dk2/basic-remote-workflow.png?raw=1)
]

>>>>

@@

From [R Studio's R Markdown Cheatsheet](https://www.rstudio.com/wp-content/uploads/2016/03/rmarkdown-cheatsheet-2.0.pdf)(images/rmd2.png)
]
]
]

@@

.pull-right[
.center[
![](https://www.dropbox.com/s/ebf8lod67dqvaoe/newrmarkdown.png?raw=1)
]
]

@@

# Structure of an R Markdown Document

@@

3. `R` chunks for data analysis, plots, figures, tables, statistics, as necessary
]

.center[
![](https://www.dropbox.com/s/og99cl4vw761m1n/rmarkdownex1.png?raw=1)
]
]
.footnote[<sup>.red[1]</sup> The one exception is for managing bibliographies, this requires an additional `.bib` file!]

@@

- You can add [additional options](https://yihui.name/knitr/options/) inside the `{braces}` after `r`, some common options:

@@

- [Markdown](https://en.wikipedia.org/wiki/Markdown) is a lightweight markup language geared towards HTML (i.e. the internet)
    - [Markup languages](https://en.wikipedia.org/wiki/Markup_language) used to add commands about how to display plain text
- Very simple and intuitive
- Write normal text as usual in any word processor
- Change font styling with tags (asterisks): 
    - `*italics text*` creates *italics text*
    - `**bold text**` creates **bold text** 
]

.pull-right[
.center[
![](https://www.dropbox.com/s/c2jq6hpqybksp6q/rmarkdownex2.png?raw=1)
]
]
---

@@

- A great resource: [Wikibooks LaTeX Mathematics chapter](https://en.wikibooks.org/wiki/LaTeX/Mathematics)

@@

- First input after the `@article` is your **citation key** (e.g. `safner2016`)
    - Whenever you want to cite this article, you'll invoke this key
]
---

@@

- Whenever you want to cite a work in your text, call up the **citation key** with `@`, like so: `@safner2016[]`, which produces @safner2016[]
    - You can customize citations, e.g.:
        - **All in parentheses**: `[@safner2016]` produces: (Safner, 2016)
        - **No parentheses**: `@safner2016` produces: Safner 2016
        - **Date only**: `[-@safner2016]` produces (2016)
        - **Adding pages or comments**: `@safner2016[p.743-744 and passim]` produces (Safner, 2016, p.743-744 and passim)

- BibTeX will automatically collect all works cited at the end and produce a **bibliography** according to a style you can choose

- We'll see more when we discuss writing your paper

@@

- Markdown files are [**plain text**](https://en.wikipedia.org/wiki/Plain_text) files and can be edited in *any* text editor
    - something as basic (and boring!) as "**Notepad**," for example
    - many good [text editors](https://en.wikipedia.org/wiki/Text_editor) out there, I like [Typora](https://www.typora.io/) or [Ulysses](https://ulysses.app/) (Mac only) for writing (and previewing) Markdown in a simple interface, with no distractions

@@

@@

- Honestly, I write **everything** in R Studio's text editor
    - Syntax highlighting
    - Actually can *run* R code, autocomplete, etc
    - Can render the markdown to an output format: html, pdf, etc.

- You can *write* R code in other text editors, but you can't *execute* them outside of *R Studio* (or the command line, but that's too advanced.) Same with actually rendering your markdown to an output (pdf, html, etc)

@@

.pull-left[
- When you are ready, you "compile" your markdown and code into an output format using:

- [`knitr`](https://yihui.name/knitr/)<sup>.red[1]</sup>, an R package that "`knit`s" your R code and markdown `.Rmd` into a `.md` file for:

- [pandoc](http://pandoc.org) is a "swiss-army knife" utility that can convert between *dozens* of document types

- All you need to do is click the `Knit` button at the top of the text editor!
]

.pull-right[
.center[
![](https://www.dropbox.com/s/kzvb8n8fpf6lstz/pandoc1.png?raw=1)

]
]
.footnote[<sup>.red[1]</sup> `knitr` also relies on the `rmarkdown` package, which will probably be installed when you first `knit.`]
---

@@

.center[
![:scale 70%](https://www.dropbox.com/s/i1ijatt56ncxi0e/github%202.png?raw=1)
]
- This project is on [GitHub](http://github.com/ryansafner/workflow)
- Click the green button, download to your computer, open `.Rproj` file in R Studio

@@

- [`GitHub`](http://github.com) is a popular (not the only!) cloud destination for these repositories

@@

.center[
![:scale 60%](https://www.dropbox.com/s/egfojonn2s54dk2/basic-remote-workflow.png?raw=1)
]
